\bf A Unitary RG-based Auxiliary Model Approach to Strongly-Correlated Electrons

### 2025-02-26 TilingProjectDraft/tilingProject.tex
updated tiled entanglement formulation and implemented corresponding numerics

\documentclass[reprint,hidelinks,onecolumn]{revtex4-2}
\usepackage{braket,amsmath,color,graphicx,bookmark,bm}
\graphicspath{{./figures/}}
\bibliographystyle{apsrev4-2}
\begin{document}

@@

\begin{figure}[!htpb]
	\centering
	\includegraphics[width=0.23\textwidth]{scattProb-3.pdf}
	\includegraphics[width=0.23\textwidth]{scattProb-5.pdf}
	\includegraphics[width=0.23\textwidth]{scattProb-6.pdf}
	\includegraphics[width=0.23\textwidth]{scattProb-7.pdf}
	\caption{Variation of \(\Gamma\) as \(W/J\) is increased from left to right and top to bottom. Each plot shows whether \(\Gamma({\bf k}\) is relevant (black) or irrelevant (white). The antinodal points are the first state to be removed from the Kondo cloud, while the node is the last to be removed.}
	\label{gamma_kxky}
\end{figure}

@@

\section{Entanglement measures}
In order to relate bipartite measures of entanglement (such as the entanglement entropy and mutual information) across the tiling algorithm, we will express the matrix elements of an arbitrary reduced density matrix (RDM) in terms of generalised correlation functions. Since these correlation functions can always be computed using the algorithm provided in the previous section, this suffices to compute the RDM using auxiliary model calculations. If the RDM is known, other bipartite measures can be easily calculated.

We consider the \(T=0\) density matrix, \(\tilde\rho_\text{gs} = \ket{\Psi}_\text{gs}\bra{{\Psi}_\text{gs}}\) of the lattice model, and a division of the total degrees of freedom (real or momentum space) into two sets \(A\) and \(B\). If basis set for the two subsystems are \(\left\{ \ket{A_i} \right\} \) and \(\left\{ B_j \right\} \), the ground state can be written as
\begin{equation}\begin{aligned}
	\ket{\Psi}_\text{gs} = \sum_{i,j}C_{i,j}\ket{A_i}\ket{B_j}~.
\end{aligned}\end{equation}
Tracing out the states of subsystem B leads to the RDM for subsystem A:
\begin{equation}\begin{aligned}
	\tilde \rho^{(A)} = \sum_k\braket{B_k | \rho | B_k} = \sum_{{i_1},{i_2}, k}C_{{i_1},k}C^*_{{i_2},k}\ket{A_{i_1}}\bra{A_{i_2}},~~\tilde \rho^{(A)}_{{i_1},{i_2}} = \sum_k C_{{i_1},k}C^*_{{i_2},k}~.
\end{aligned}\end{equation}
We will now show that the product of coefficients \(C_{i,k}C^*_{j,k}\) can also be obtained through expectation value calculations of product operators in the subspace of \(A\). Consider the following transition operator:
\begin{equation}\begin{aligned}
	P^{(A)}_{i_1,i_2} = \ket{A_{i_1}}\bra{A_{i_2}}~.
\end{aligned}\end{equation}
The RDM elements are expectation values of this operator in the ground state:
\begin{equation}\begin{aligned}
	\braket{P^{(A)}_{i_1,i_2}} = \sum_k C_{{i_1},k}C^*_{{i_2},k} = \tilde \rho^{(A)}_{{i_1},{i_2}}~.
\end{aligned}\end{equation}

We now argue that these expectation values are simply correlation functions corresponding to products of local operators. If the subsystem \(A\) is composed of \(N_A\) one-particle Hilbert spaces \(A^{(\nu)}\), any basis state \(\ket{A_{i}}\) can be written as the product of one-particle basis states: 
\begin{equation}\begin{aligned}
	\ket{A_i} = \otimes_{\nu=1}^{N_A}\ket{s_\nu(A_i)}~,
\end{aligned}\end{equation}
where \(s_\nu(A_i)\) can be 0 or 1 and represents the configuration (occupied/unoccupied) of the local degree of freedom \(A^{(\nu)}\) in the basis state \(\ket{A_i}\), and \(\ket{s_\nu(A_i)}\) is the state of \(A^{(\nu)}\) in that configuration. The transition operator \(P^{(A)}_{i_1,i_2}\) can therefore be written as 
\begin{equation}\begin{aligned}
	P^{(A)}_{i_1,i_2} = \ket{s_1(A_{i_1})}\otimes\ket{s_2(A_{i_1})} \ldots\otimes\ket{s_{N_A}(A_{i_1})}\bra{s_{N_A}(A_{i_2})}\otimes\ldots\bra{s_2(A_{i_2})}\otimes\bra{s_1(A_{i_2})}~.
\end{aligned}\end{equation}
This can be converted into a second-quantised operator, by making the substitution:
\begin{equation}\begin{aligned}
	\ket{s} \to 
	\begin{cases}
		c^\dagger,~\text{if }s=1,\\
		1-n,~\text{if }s=0,	
	\end{cases}~.
\end{aligned}\end{equation}
As an example, for a two-qubit subspace \(A\), the transition operator \(P^{(A)}_{00,01}\) between the states \(i_1=00\) (both unoccupied) and \(i_2=01\) (only one occupied) can be expressed as
\begin{equation}\begin{aligned}
	P^{(A)}_{00,01} = \ket{0}_1\ket{0}_2\bra{1}_2\bra{0}_1 = (1-n_1)(1-n_2)(1-n_2)c_1 = c_1(1-n_2)~.
\end{aligned}\end{equation}
This means the RDM element for this subspace, at the location \((00, 01)\), is equal to the ground state expectation value of the operator \(c_1(1-n_2)\). The process of calculating such expectation values, within the present algorithm, has already been explained in Sec.~\ref{staticCorr}. Note that this allows the computation of reduced density matrix for real-space as well as momentum-space partitioning.

Other measures like mutual information can be obtained by using these elementary measures.

\section{Metal-insulator transition in the Lattice Model}
\subsection{Momentum-space picture etc etc}
In order to understand the nature of the transition in momentum space, we employ the expression obtained in the previous sections and compute the spectral function, self-energy and correlation functions in the top right quadrant of the first Brillouin zone. We first show the \(k-\)space spectral functions near zero frequency, in Fig.~\ref{kspace_specfunc}.
\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.23\textwidth]{tiledSF-1.pdf}
	\includegraphics[width=0.23\textwidth]{tiledSF-2.pdf}
	\caption{tiledSF.pdf}
	\label{fig:tiledSF-pdf}
\end{figure}
% \begin{figure*}[htpb]
% 	\centering
% 	\includegraphics[width=\textwidth]{charge-p-p_37_-2.0_-1.7297_-0.0_0.1.pdf}
% \end{figure*}
% \begin{figure*}[htpb]
% 	\centering
% 	\includegraphics[width=\textwidth]{vne-p-p_37_-2.0_-1.7297_-0.0_0.1.pdf}
% \end{figure*}

@@

The impurity retarded Green's function (assuming the Hamiltonian to be time-independent, which it is) is defined as

@@

	G_{dd}^\sigma(t) = -i\theta(t) \left<\left\{ \mathcal{O}_{\sigma}(t),\mathcal{O}^\dagger_{\sigma} \right\}  \right>

@@

where the average \(\left< \right>\) is over a canonical ensemble at temperature \(T\), and \(\mathcal{O}_\sigma = c_{d\sigma} + S_d^- c_{0\overline\sigma} + S_d^z c_{0\sigma}\) is the excitation whose spectral function we are interested in. The excitations defined in \(\mathcal{O}\) incorporates both single-particle excitations brought about by the hybridisation as well as two-particle spin excitations brought about by the spin-exchange term. What follows is a standard calculation where we write the Green's function in the Lehmann-Kallen representation. The ensemble average for an arbitrary operator \(\hat M\) can be written in terms of the exact eigenstates of the fixed point Hamiltonian:

@@

	H^* \ket{n} &= E_n^* \ket{n}, ~ ~ ~\left<\hat M \right> &\equiv \frac{1}{Z}\sum_n \bra{n} \hat M \ket{n} e^{-\beta E_n^*}

@@

where \(Z = \sum_n e^{-\beta E_n^*}\) is the fixed point partition function and \(\left\{ \ket{n} \right\} \) is the set of eigenfunctions of the fixed point Hamiltonian. We can therefore write

@@

	G_{dd}^\sigma = -i\theta(t)\frac{1}{Z}\sum_{m,n}e^{i\left( E_m - E_n \right)  t}||\bra{m}\mathcal{O}_{\sigma}\ket{n}||^2\left( e^{-\beta E_m} + e^{-\beta E_n}\right)\\

@@

	G_{d d}^\sigma(\omega) &= \int_{-\infty}^\infty dt e^{i \omega t}G_{d d}^\sigma(t) \\
			       &= \frac{1}{Z}\sum_{m,n}||\bra{m}\mathcal{O}_{\sigma}\ket{n}||^2\left( e^{-\beta E_m} + e^{-\beta E_n}\right)\left(-i\right)\int_{-\infty}^\infty dt \theta(t)e^{i\left( \omega + E_m - E_n \right)t}

@@

	\theta(t) = \frac{1}{2\pi i}\lim_{\eta \to 0^+} \int_{-\infty}^\infty \frac{1}{x- i\eta}e^{ixt}dx

@@

With this definition, the integral in \(G_{dd}^\sigma(\omega)\) becomes

@@

	\left(-i\right)\int_{-\infty}^\infty dt \theta(t)e^{i\left( \omega + E_m - E_n \right)t} &= \left(-i\right)\frac{1}{2\pi i}\lim_{\eta \to 0^+} \int_{-\infty}^\infty dx\frac{1}{x- i\eta}\int_{-\infty}^\infty dt e^{i\left( \omega + E_m - E_n + x\right)t} \\
									     &=\left(-i\right)\frac{1}{2\pi i}\lim_{\eta \to 0^+} \int_{-\infty}^\infty dx\frac{1}{x- i\eta} 2\pi \delta\left( \omega + E_m - E_n + x\right) \\
									     &=\left(-i\right)\frac{1}{i}\lim_{\eta \to 0^+} \frac{-1}{\omega + E_m - E_n- i\eta} \\
									     &=\frac{1}{\omega + E_m - E_n} \\

@@

	G_{d d}^\sigma(\omega) = \frac{1}{Z}\sum_{m,n}||\bra{m}\mathcal{O}_{\sigma}\ket{n}||^2\left( e^{-\beta E_m} + e^{-\beta E_n}\right)\frac{1}{\omega + E_m - E_n}

@@

The zero temperature Green's function is obtained by taking the limit of \(\beta \to \infty\). In both the partition function as well as inside the summation, the only term that will survive is the exponential of the ground state energy \(E_0\).
\begin{equation*}\begin{aligned}
	Z \equiv \sum_m e^{-\beta E_m} \implies \lim_{\beta \to \infty}Z = d_0 e^{-\beta E_0}, && E_0 \equiv \text{min}\left\{ E_n \right\} 
\end{aligned}\end{equation*}
where \(d_0\) is the degeneracy of the ground state. The Greens function then simplifies to

@@

	\label{Gf_lk}
	G_{d d}^\sigma(\omega, \beta \to \infty) &= \frac{1}{d_0e^{-\beta E_0}}\sum_{m,n}||\bra{m}\mathcal{O}_{\sigma}\ket{n}||^2\left[e^{-\beta E_m}\delta_{E_m, E_0} + e^{-\beta E_n}\delta_{E_n, E_0}\right]\frac{1}{\omega + E_m - E_n}\\
						 &= \frac{1}{d_0}\sum_{n,0}\left[||\bra{0}\mathcal{O}_{\sigma}\ket{n}||^2\frac{1}{\omega + E_0 - E_n} + ||\bra{n}\mathcal{O}_{\sigma}\ket{0}||^2\frac{1}{\omega - E_0 + E_n}\right]\\

@@

The label 0 sums over all states \(\ket{0}\) with energy \(E_0\). The spectral function is the imaginary part of this Green's function. To extract the imaginary part, we insert an infinitesimal imaginary part in the denominator:

@@

	G_{d d}^\sigma(\omega, \eta) = \frac{1}{d_0}\lim_{\eta \to 0^-}\sum_{n,0}\left[||\bra{0}\mathcal{O}_{\sigma}\ket{n}||^2\frac{1}{\omega + E_0 - E_n + i\eta} + ||\bra{n}\mathcal{O}_{\sigma}\ket{0}||^2\frac{1}{\omega - E_0 + E_n + i\eta}\right]\\

@@

The spectral function at zero temperature can then be written as

@@

	\mathcal{A(\omega)} &= -\frac{1}{\pi}\text{ Im }\left[G_{dd}^\sigma\left( \omega \right) \right] \\
			    &= \frac{1}{d_0}\frac{1}{\pi}\text{ Im }\left[\lim_{\eta \to 0^-}\sum_{n,0}\left(\frac{-i\eta ||\bra{0}\mathcal{O}_{\sigma}\ket{n}||^2}{\left(\omega + E_0 - E_n\right)^2 + \eta^2} + \frac{-i\eta ||\bra{n}\mathcal{O}_{\sigma}\ket{0}||^2}{\left(\omega - E_0 + E_n\right)^2 + \eta^2}\right)\right]\\
			    &= \frac{1}{d_0}\frac{1}{\pi}\sum_{n,0}\left[||\bra{0}\mathcal{O}_{\sigma}\ket{n}||^2\pi\delta\left(\omega + E_0 - E_n\right) + ||\bra{n}\mathcal{O}_{\sigma}\ket{0}||^2\pi\delta\left(\omega - E_0 + E_n\right)\right]\\

>>>>

\documentclass[reprint,hidelinks,onecolumn]{revtex4-2}
\usepackage{braket,amsmath,color,graphicx,bookmark,bm,amssymb}
\graphicspath{{./figures/}}
\bibliographystyle{apsrev4-2}
\begin{document}

@@

@@

\section{Entanglement}
We will now describe the prescription of calculating entanglement measures of the lattice model from within our auxiliary model treatment. In this section, we are interested mainly in two such measures, the entanglement entropy and the mutual information. Given a pure state \(\ket{\Psi}\) describing the complete system, the entanglement entropy \(S_\text{EE}(\nu)\) of a subsystem \(\nu\) quantifies the entanglement of \(\nu\) with the rest of the subsystem, and is defined as
\begin{equation}\begin{aligned}
	S_\text{EE}(\nu) = -\text{Tr}\left[\rho(\nu)\log \rho(\nu)\right],\quad \rho(\nu) = \text{Tr}_\nu\left[\ket{\Psi}\bra{\Psi}\right] 
\end{aligned}\end{equation}
where \(\text{Tr}\left[\cdot\right] \) is the trace operator, and \(\rho(\nu)\) is the reduced density matrix for the subsystem \(\nu\) obtained by taking the partial trace \(\text{Tr}_\nu\) (over the states of \(\nu\)) of the full density matrix \(\rho = \ket{\Psi}\bra{\Psi}\). If the subsystem \(\nu\) describes local regions in real space (or states in \(k-\)space), we might be interested in the entanglement between two such subsystems \(\nu_1\) and \(\nu_2\). The correct measure to quantify such entanglement is the mutual information:
\begin{equation}\begin{aligned}
	I_2(\nu_1,\nu_2) = S_\text{EE}(\nu_1) + S_\text{EE}(\nu_2) - S_\text{EE}(\nu_1 \cup \nu_2)~,
\end{aligned}\end{equation}
where \(\nu_1 \cup \nu_2\) is a larger subsystem formed by combining \(\nu_1\) and \(\nu_2\).

\subsection{Real-space entanglement}
Real-space entanglement measures can be used to probe delocalisation-localisation transitions. The simplest such measure is the entanglement of a local mode. Since the local entanglement entropy will be uniform at each lattice site for a system with translation invariance, it suffices to calculate the real-space averaged entanglement entropy \(S_\text{EE}^\text{loc} = \frac{1}{N}\sum_{{\bf r}}S_\text{EE}\left( {\bf r} \right) \). By visualising the lattice model as a superposition of auxiliary models placed at various sites (eq.~\ref{tilingPrescriptionFinal}), the real space average of the lattice model can be thought of as an average over sites of a particular impurity model, and then a second average over all the impurity models. But since all all impurity models are equivalent to each other, the second average is redundant. Secondly, all correlations must derive from the impurity site, which can be formally encoded by subtracting, from this average, the corresponding contribution obtained in the absence of the impurity site. In total, the lattice-auxiliary model relation for the local entanglement entropy can be written as
\begin{equation}\begin{aligned}
	S_\text{EE}^\text{loc} = \frac{1}{N}\sum_{{\bf r}}\left[S_\text{EE}({\bf r}_d + {\bf r}) - S_\text{EE}^{(0)}({\bf r}_d + {\bf r})\right] ~,
\end{aligned}\end{equation}
where \({\bf r}_d\) is the impurity site position, \({\bf r}\) is the distance of a conduction bath site from the impurity site (can be zero), and \(S_\text{EE}({\bf r}_d + {\bf r})\) is the entanglement entropy, calculated within the impurity model, at the locatiom \({\bf r}_d + {\bf r}\). \(S^{(0)}_\text{EE}({\bf r}_d + {\bf r})\) is the same entanglement entropy, but calculated for an impurity model but with vanishing impurity-bath hybridisation and bath interaction.

\subsection{Momentum-space entanglement}
Entanglement measures in \(k-\)space can provide valuable information regarding the Fermi surface structure and the nature of gapless excitations proximate to it. The most elementary measure is the entanglement entropy \(S_\text{EE}({\bf q})\) of a single excitation carrying momentum \({\bf q}\). Unlike \(S_\text{EE}^\text{loc}\), this does not involve an average, and involves a single computation:
\begin{equation}\begin{aligned}
	S_\text{EE}^\text{loc} = S_\text{EE}({\bf q}) - S_\text{EE}^{(0)}({\bf q})~.
\end{aligned}\end{equation}
It is possible to improve this by considering inter-auxiliary model contributions in the reduced density matrix \(\rho({\bf q})\). On the lattice model, \(\rho({\bf q})\) is defined as
\begin{equation}\begin{aligned}\label{SEE-k}
	\rho({\bf q}) = \text{Tr}_{\bf q}\left[\ket{\Psi_\text{gs}}\bra{\Psi_\text{gs}}\right] ~,
\end{aligned}\end{equation}
where \(\text{Tr}_{\bf q}\left[\cdot\right]\) is the partial trace over the Hilbert space of \({\bf q}\), and \(\ket{\Psi_\text{gs}}\) is the tiled ground state. Using eq.~\ref{groundstateProposal}, we can write the ground state in terms of those of the auxiliary model. These leads to two classes of terms, one purely within a single auxiliary model ground state \(\ket{\psi_\text{gs}}\), and the other involving transitions across auxiliary models:
\begin{equation}\begin{aligned}
	\rho({\bf q}) = \text{Tr}_{\bf q}\left[\ket{\psi_\text{gs}({\bf r}_d)}\bra{\psi_\text{gs}({\bf r}_d)}\right] + \frac{1}{N}\text{Tr}_{\bf q}\sum_{{\bf r}}\left[\ket{\psi_\text{gs}({\bf r}_d)}\bra{\psi_\text{gs}({\bf r}_d + {\bf r})}\right]~,
\end{aligned}\end{equation}
where \(\ket{\psi_\text{gs}({\bf r}_d)}\) is a reference auxiliary model ground state, and \(\ket{\psi_\text{gs}({\bf r}_d + {\bf r})}\) sums over other auxiliary models at increasing distances from this reference model. The first term leads to the expression in eq.~\ref{SEE-k}, while the second term can be used to improve this estimate. Away from any quantum critical points, the localised nature of the impurity-bath hybridisation ensures that impurity correlations decay exponentially away from the transition. This ensures that the second term is very small away from a critical point. We therefore restrict ourselves to just the first term in the present work.

\begin{figure}[!htpb]
	\centering
	\includegraphics[width=0.23\textwidth]{scattProb-3.pdf}
	\includegraphics[width=0.23\textwidth]{scattProb-5.pdf}
	\includegraphics[width=0.23\textwidth]{scattProb-6.pdf}
	\includegraphics[width=0.23\textwidth]{scattProb-7.pdf}
	\caption{Variation of \(\Gamma\) as \(W/J\) is increased from left to right and top to bottom. Each plot shows whether \(\Gamma({\bf k})\) is relevant (white) or irrelevant (black). The antinodal points are the first state to be removed from the Kondo cloud, while the node is the last to be removed.}
	\label{gamma_kxky}
\end{figure}

\section{Journey Through the Pseudogap on The Lattice model}
We have already seen from the lattice-embedded impurity model that the impurity phase transition is anisotropic in \(k-\)space. This has consequences for the metal-insulator transition of the lattice model. We now describe, in terms of correlations, how the Fermi surface of the lattice model is destabilised when the system is tuned through the pseudogap. For this, we employ the relations we derived between correlations on the impurity model and those on the lattice model.

\subsection{One-particle \(k-\)space spectral function}
\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.23\textwidth]{kspaceDOS-1.pdf}
	\includegraphics[width=0.23\textwidth]{kspaceDOS-2.pdf}
	\includegraphics[width=0.23\textwidth]{kspaceDOS-3.pdf}
	\includegraphics[width=0.23\textwidth]{kspaceDOS-4.pdf}
	\caption{One-particle \(k-\)space spectral function. In the absence of bath interactions, all points on Fermi surface are gapless (first panel). Proceeding through the pseudogap results in the gapping out of Fermi surface points starting from the antinode (second panel) and approaching the node (third panel), until finally only the node remains at the critical point (fourth panel).}
	\label{fig:tiledSF-pdf}
\end{figure}

\subsection{Momentum-space spin-correlations}
We first look at static spin correlations \(\chi_s(k_1, k_2) = \frac{1}{2}\left<S_{k_1}^+ S_{k_2}^- + \text{h.c.} \right>\). These processes correspond to the gapless excitations of the Fermi liquid phase. We find that 
\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.24\textwidth]{tiledSF-1.pdf}
	\includegraphics[width=0.24\textwidth]{tiledSF-2.pdf}
	\includegraphics[width=0.24\textwidth]{tiledSF-3.pdf}
	\includegraphics[width=0.24\textwidth]{tiledSF-4.pdf}
	\caption{Spin-flip correlations \(\left<S(k_N)^+, S({\bf k})^-\right>\) between the nodal point \(k_N = \left( -\pi/2, -\pi/2\right)\) and an arbitrary \(k-\)point on the Fermi surface. In the absence of bath interaction (first panel), the correlations are somewhat uniformly distributed along the Fermi surface. As we enter the pseudogap (second panel), the spin-correlations near the antinode vanish, indicating that they have been removed from the metallic excitations. Passing through the pseudogap involves in the extensions of the gapped region (third panel), until finally only the node remains (fourth panel).}
	\label{tiledSF}
\end{figure}

@@

\subsection{Spectral representation of Greens function}
The impurity retarded Green's function is defined as

@@

	G_{d\sigma}(t) = -i\theta(t) \left<\left\{ \mathcal{O}_{\sigma}(t),\mathcal{O}^\dagger_{\sigma} \right\}  \right>

@@

where the average \(\left< \right>\) is over a canonical ensemble at temperature \(T\), and \(\mathcal{O}_\sigma\) is the excitation whose spectral function we are interested in. What follows is a standard calculation where we write the Green's function in the spectral representation. The ensemble average for the operator \(\mathcal{O}_\sigma\) can be written in terms of the exact eigenstates of the Hamiltonian:

@@

	H \ket{n} &= E_n \ket{n}, \quad\left<\mathcal{O}_\sigma\right> &\equiv \frac{1}{Z}\sum_n \braket{n | \mathcal{O}_\sigma | n} e^{-\beta E_n}

@@

where \(Z = \sum_n e^{-\beta E_n}\) is the partition function and \(\left\{ \ket{n} \right\} \) is the set of eigenfunctions of the Hamiltonian. We can therefore write

@@

	G_{d\sigma} = -i\theta(t)\frac{1}{Z}\sum_{m,n}e^{i\left( E_m - E_n \right)  t}||\bra{m}\mathcal{O}_{\sigma}\ket{n}||^2\left( e^{-\beta E_m} + e^{-\beta E_n}\right)\\

@@

	G_{d d}^\sigma(\omega) &= \int_{-\infty}^\infty dt e^{i \left(\omega \right)t}G_{d d}^\sigma(t) \\
			       &= \frac{1}{Z}\sum_{m,n}||\bra{m}\mathcal{O}_{\sigma}\ket{n}||^2\left( e^{-\beta E_m} + e^{-\beta E_n}\right)\left(-i\right)\int_{-\infty}^\infty dt \theta(t)e^{i\left(\omega + E_m - E_n \right)t}

@@

	\theta(t) = -\frac{1}{2\pi i}\lim_{\eta \to 0^+} \int_{-\infty}^\infty \frac{1}{x + i\eta}e^{-ixt}dx

@@

With this definition, the integral in \(G_{d\sigma}(\omega)\) becomes

@@

	\left(-i\right)\int_{-\infty}^\infty dt \theta(t)e^{i\left( \omega + i0^+ + E_m - E_n \right)t} &= \left(-i\right)\frac{-1}{2\pi i}\lim_{\eta \to 0^+} \int_{-\infty}^\infty dx\frac{1}{x+ i\eta}\int_{-\infty}^\infty dt e^{i\left( \omega + i0^+ + E_m - E_n - x\right)t} \\
									     &=\frac{1}{2\pi}\lim_{\eta \to 0^+} \int_{-\infty}^\infty dx\frac{1}{x+ i\eta} 2\pi \delta\left( \omega + i0^+ + E_m - E_n - x\right) \\
									     &=\frac{1}{\omega + E_m - E_n + i0^+}~.

@@

	G_{d d}^\sigma(\omega) = \frac{1}{Z}\sum_{m,n}||\bra{m}\mathcal{O}_{\sigma}\ket{n}||^2\left( e^{-\beta E_m} + e^{-\beta E_n}\right)\frac{1}{\omega + E_m - E_n + i0^+}

@@

The infinitesimally positive imaginary part in the denominator shifts the poles onto the lower half of the complex plane, leaving it analytic in the upper half: this is necessary to make the retarded Greens function causal.

\subsection{Real and imaginary parts - The Sokhotski-Plemelj theorem}
In order to write down the real and imaginary parts of the Greens function, we first prove the {\it Sokhotski-Plemelj} formula:
\begin{equation}\begin{aligned}\label{SokhotskiPlemelj}
	\lim_{\eta \to 0^+} \frac{1}{x + i \eta} = P\left(\frac{1}{x}\right) - i \pi \delta(x)~.
\end{aligned}\end{equation}
where \(P(f(x))\) is the Cauchy principal value, defined as 
\begin{equation}\begin{aligned}
	P\left[\int_{-\infty}^{\infty}\frac{dx}{x}\right] = \int_{-\infty}^{0^-}\frac{dx}{x} + \int_{0^+}^{\infty}\frac{dx}{x}~.
\end{aligned}\end{equation}

To prove the above identity, we integrate the left-hand side using a test function \(f(x)\):
\begin{equation}\begin{aligned}\label{twoIntegrals}
	\lim_{\eta \to 0^+}\int_{-\infty}^{\infty} \text{dx}~\frac{f(x)}{x + i \eta} &= \lim_{\eta \to 0^+}\int_{-\infty}^{\infty} \text{dx}~\frac{f(x)\left( x - i\eta \right) }{x^2 + \eta^2} \\
																				 &= \lim_{\eta \to 0^+}\int_{-\infty}^{\infty} \text{dx}~f(x)\left[\frac{x}{x^2 + \eta^2} - i\eta\frac{1}{x^2 + \eta^2}\right]
\end{aligned}\end{equation}
The first integral can be split into three parts:
\begin{equation}\begin{aligned}
	\lim_{\eta \to 0^+}\int_{-\infty}^{\infty} \text{dx}~f(x)\frac{x}{x^2 + \eta^2} &= \lim_{\eta \to 0^+}\lim_{\varepsilon_\to 0^+}\left[\int_{-\infty}^{-\varepsilon} \text{dx}~\frac{f(x) x}{x^2 + \eta^2} + \int_{\varepsilon}^{\infty} \text{dx}~\frac{f(x)x}{x^2 + \eta^2} + \int_{-\varepsilon}^{\varepsilon} \text{dx}~\frac{f(x)x}{x^2 + \eta^2}\right]
\end{aligned}\end{equation}
For the first two parts, it is safe to take the limit of \(\eta\), because \(x\) is always non-vanishing there. For the third part, we can approximate \(f(x)\) as \(f(0)\) in the neighbourhood \(x \in \left[-\varepsilon, \varepsilon\right], \varepsilon \to 0 \). This leaves an odd function \(x/(x^2 + \eta^2)\) as the integrand, being integrated over a symmetric range. The third term therefore vanishes. In total, we get
\begin{equation}\begin{aligned}\label{integral1}
	\lim_{\eta \to 0^+}\int_{-\infty}^{\infty} \text{dx}~f(x)\frac{x}{x^2 + \eta^2} &= \int_{-\infty}^{-\varepsilon} \text{dx}~\frac{f(x)}{x} + \int_{\varepsilon}^{\infty} \text{dx}~\frac{f(x)}{x}  = P\left(\int_{-\infty}^{\infty} \text{dx}~\frac{f(x)}{x}\right) ~.
\end{aligned}\end{equation}
To evaluate the second integral of eq.~\ref{twoIntegrals}, we note that the only region in which the integrand is non-zero is when \(|x| \simeq 0\); there, we again approximate \(f(x)\) as \(f(0)\). The remaining integral can then be evaluated easily:
\begin{equation}\begin{aligned}\label{integral2}
	-i\lim_{\eta \to 0^+}\eta\int_{-\infty}^{\infty} \text{dx}~\frac{f(x)}{x^2 + \eta^2} = -i \lim_{\eta \to 0^+}\eta f(0)\int_{-\infty}^{\infty} \frac{\text{dx}}{x^2 + \eta^2} = -i \pi f(0),
\end{aligned}\end{equation}
where we used \(\int \frac{dx}{x^2 + \eta^2} = \frac{1}{\eta} \arctan\left( x/\eta \right) \). Combining eqs.~\ref{twoIntegrals},\ref{integral1} and \ref{integral2}, we get
\begin{equation}\begin{aligned}
	\lim_{\eta \to 0^+}\int_{-\infty}^{\infty} \text{dx}~\frac{f(x)}{x + i \eta} = P\left(\int_{-\infty}^{\infty} \text{dx}~\frac{f(x)}{x}\right) - i\pi \int_{-\infty}^{\infty} \text{dx}~f(x)\delta(x)~.
\end{aligned}\end{equation}
This proves eq.~\ref{SokhotskiPlemelj}.

The Sokhotski-Plemelj formula allows us to split the spectral representation into a real and an imaginary part:
\begin{equation}\begin{aligned}
	G_{d\sigma}^\prime = \frac{1}{Z}\sum_{m,n}||\bra{m}\mathcal{O}_{\sigma}\ket{n}||^2\left( e^{-\beta E_m} + e^{-\beta E_n}\right)P\left(\frac{1}{\omega + E_m - E_n}\right)\\
	G_{d\sigma}^{\prime\prime} = -\pi\frac{1}{Z}\sum_{m,n}||\bra{m}\mathcal{O}_{\sigma}\ket{n}||^2\left( e^{-\beta E_m} + e^{-\beta E_n}\right)\delta\left(\omega + E_m - E_n\right)~.
\end{aligned}\end{equation}
The spectral function, specifically, is defined as follows:
\begin{equation}\begin{aligned}
	A_{d\sigma}(\omega) = -\frac{1}{\pi}G_{d\sigma}^\prime = \frac{1}{Z}\sum_{m,n}||\bra{m}\mathcal{O}_{\sigma}\ket{n}||^2\left( e^{-\beta E_m} + e^{-\beta E_n}\right)\delta\left(\omega + E_m - E_n\right)
\end{aligned}\end{equation}

\subsection{Zero temperature spectral function}
We specialise to zero temperature by taking the limit of \(\beta \to \infty\). In both the partition function as well as inside the summation, the only term that will survive is the exponential of the ground state energy \(E_\text{GS}\).
\begin{equation*}\begin{aligned}
	Z \equiv \sum_m e^{-\beta E_m} \implies \lim_{\beta \to \infty}Z = d_\text{GS} e^{-\beta E_\text{GS}}, && E_\text{GS} \equiv \text{min}\left\{ E_n \right\} 
\end{aligned}\end{equation*}
where \(d_\text{GS}\) is the degeneracy of the ground state. The spectral function then simplifies to
\begin{equation}\begin{aligned}
	\label{Gf_lk}
	A_{d\sigma} &= \frac{1}{d_\text{GS}e^{-\beta E_\text{GS}}}\sum_{m,n}||\bra{m}\mathcal{O}_{\sigma}\ket{n}||^2\left[e^{-\beta E_m}\delta_{E_m, E_\text{GS}} + e^{-\beta E_n}\delta_{E_n, E_\text{GS}}\right]\delta\left(\omega + E_m - E_n\right)\\
											 &= \frac{1}{d_\text{GS}}\sum_{n,n_\text{GS}}\left[||\bra{n_\text{GS}}\mathcal{O}_{\sigma}\ket{n}||^2\delta\left(\omega + E_\text{GS} - E_n\right) + ||\bra{n}\mathcal{O}_{\sigma}\ket{n_\text{GS}}||^2\delta\left(\omega - E_\text{GS} + E_n\right)\right]\\
\end{aligned}\end{equation}
The label \(n_\text{GS}\) sums over all states \(\ket{n_\text{GS}}\) with energy \(E_\text{GS}\). In practise, we evaluate this expression by replacing the formal Dirac delta functions with "nascent" ones, such as a Lorentzian function with a sufficiently sharp peak.

\subsection{Reconstructing full Green function from spectral function: Kramers-Kronig relations}
As mentioned earlier, the retarded Greens function \(G_R(\omega + i 0^+)\) has a complex pole in the lower half of the complex plane. In the following, the frequency \(\omega\) is assumed to be complex. Consider the integral

@@

	I(\omega) = \oint_C d\omega^\prime \frac{G_R(\omega^\prime)}{\omega^\prime - \omega},~\omega \in \mathbb{R}~.

@@

The contour \(C\) encloses the entirety of the upper half of the complex plane as well as almost all points of the real line but avoids the pole at \(\omega^\prime = \omega\) by forming a semicircle about that point that extends into the upper half of the plane. Since the integrand is completely analytic in the region enclosed by the contour (recall again that \(G_R\) is analytic in the upper half), the integral evaluates to zero (by Cauchy's theorem):
\begin{equation}\begin{aligned}\label{zerointegral}
	I(\omega) = 0~.
One can also evaluate the integral along each part of the contour. The semicircular part contributes zero: this is because while the length of the semicircular arc goes as \(|\omega^\prime|\), the integrand vanishes faster than \(1/|\omega^\prime|\) as \(|\omega^\prime| \to \infty\) (this assumes that the Green function vanishes in the limit of \(\omega \to \infty\)). The part of the contour along the real axis can now be evaluated. The two parts of the contour on either side of the pole contribute

@@

	\int_{-\infty}^{\omega - 0^+}d\omega^\prime~\frac{G_R(\omega^\prime)}{\omega^\prime - \omega} + \int_{\omega + 0^+}^{\infty}d\omega^\prime~\frac{G_R(\omega^\prime)}{\omega^\prime - \omega} = P\int_{-\infty}^{\infty}d\omega^\prime~\frac{G_R(\omega^\prime)}{\omega^\prime - \omega}

@@

To evaluate the part of the integral along the small semicircular around \(\omega^\prime = \omega\) (that extends into the upper half), we argue that this contribution should be equal to that obtained by taking the semicircle that goes instead into the lower half but circulates in the opposite direction. Each of these contributions should, in turn, be equal to that obtained from joining these two contours. But joining these contours leads to a closed integral about the pole at \(\omega^\prime = \omega\), which is equal to \(2\pi G_R(\omega)\).

@@

	\int_\text{upper} d\omega^\prime~\frac{G_R(\omega^\prime)}{\omega^\prime - \omega} = \int_\text{lower} d\omega^\prime~\frac{G_R(\omega^\prime)}{\omega^\prime - \omega} = \frac{1}{2}\oint_{C(\omega)} d\omega^\prime~\frac{G_R(\omega^\prime)}{\omega^\prime - \omega} = i\pi G_R(\omega)~.
